---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 3 Notes
### Created by Christopher Kinson


***


## Things Covered in this Week's Notes

- **An Introduction to R** *Chapter 3 Objects, their modes and attributes*

- Object Arrangement

- **An Introduction to R** *Chapter 4 Ordered and unordered factors*


***


### ICYMI: [**An Introduction to R**](https://cran.r-project.org/doc/manuals/r-release/R-intro.pdf) by Venables, Smith and the R Core Team is one of the reference textbooks in the syllabus.


***


## **An Introduction to R** *Chapter 3 Objects, their modes and attributes*

**Atomic** objects contain values all of the same type - character, logical, numeric. Vectors, matrices, and arrays are atomic structures. **Recursive** objects contain values of different types. Lists, functions, and expressions are recursive objects. Every defined object has a mode and a length.

```{r note1}
mode(letters)
length(letters)
```

We can also check the **structure** of an object which includes the mode or type and the first few elements of each of the object's dimensions.
```{r}
str(letters)
```


Some common modes include numeric, logical, character, list, matrix, array, factor, and data.frame. There are other ways to check the mode of an object that would return a logical value of `TRUE` or `FALSE`. 
```{r}
is.atomic(letters)
is.numeric(letters)
is.logical(letters)
is.character(letters)
is.list(letters)
is.matrix(letters)
is.array(letters)
is.factor(letters)
is.data.frame(letters)
```

An object's **mode** can be changed via coercion. Coercion is the act of changing the mode of an object. R allows for this to occur for many modes and objects even when it doesn't make sense. Coercion can happen by using the `as.newmode(object)`. See below for examples.

```{r note2}
x <- 1:20
x <- x[1:19]
xx <- as.character(x)
xx
as.numeric(xx)
```

An object's **length** can be changed via reassignment.
```{r}
length(xx) <- 10
```

## Object Arrangement

Objects can be sorted or arranged in an order that we have control over. Objects can be arranged with the `sort()` function or the `order()` function. The `sort()` function arranges the object displaying the raw values of the object in its new order. Whereas the `order()` function arranges the object displaying the corresponding indices; not the raw values. We can combine these functions with the concept of indexing (week 2).
```{r}
iris_SL <- iris$Sepal.Length[1:20]
sort(iris_SL)
sort(iris_SL, decreasing = TRUE) #decreasing order
order(iris_SL)
iris_SL[order(iris_SL)]
iris_SL[order(iris_SL, decreasing = TRUE)] #decreasing order
```


## **An Introduction to R** *Chapter 4 Ordered and unordered factors*

Factors are vectors specially designed for categorical data. Their levels are the categories within. Inside a factor, its levels may be ordered or unordered. When we want to see the levels, they are usually in alphanumeric order.

```{r note3}
data(iris)
iris$Species
is.factor(iris$Species)
levels(iris$Species)
factor(round(iris$Sepal.Length,0))
```

Some times, factors are not preferred. In previous versions of R (before 4.0.0), data frames automatically coerced any column with characters or strings as factor type. We can imagine that this default behavior would not always be welcome. And the `tidyverse` collection of packages made sure to change this default behavior. We'll discuss more about the tidyverse functionality later in the semester. *If you don't have it installed already, do install the `tidyverse` package; otherwise the chunk won't run properly.*

```{r}
#if your version of R is outdated, then you will see differences in the following results when you execute this chunk
str(read.csv("https://uofi.box.com/shared/static/ko6uqp336nruywnormrp6euo9qrraigs.csv"))
library(tidyverse)
str(read_csv("https://uofi.box.com/shared/static/ko6uqp336nruywnormrp6euo9qrraigs.csv"))
```

