---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 6 Practice Problems Solutions
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#1** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the Three Sigma Rule given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers and the index location of the outlier (i.e. if the first value of x is an outlier then your function returns 1 and the value at x[1]).

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 

```{r}
three_sigma_rule <- function(data){
  #data[abs(data-mean(data))>3*sd(data)]
  data.frame(outlier_value=data[abs(data-mean(data))>3*sd(data)], outlier_position=which(abs(data-mean(data))>3*sd(data)))
}
x <- iris$Sepal.Width
three_sigma_rule(x)
```


**#2** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the 1.5*IQR Rule given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers.

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

```{r}
iqr_rule <- function(data){
  data.frame(outlier_value=data[data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)], outlier_position=which(data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)))
}
x <- iris$Sepal.Width
iqr_rule(x)
```


**#3** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the Hampel Identifier given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers.

- Hampel Identifier $|x-\tilde{x}|> 3 \cdot \tilde{\sigma}$  
      - where $\tilde{x}$ is the median of $x$  
      - where $\tilde{\sigma}=1.4826 \cdot \textrm{median}\{|x-\tilde{x}|\}$ is the median of the absolute deviation from the median (MADM or MAD) scale estimate

```{r}
hampel_identifier <- function(data){
  data.frame( outlier_value=data[abs(data-median(data))>3*1.4826*median(data-median(data))], outlier_position=which(abs(data-median(data))>3*1.4826*median(data-median(data))) )
}
x <- iris$Sepal.Width
hampel_identifier(x)

```

