---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 6 Practice Problems
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#1** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the Three Sigma Rule given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers.

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 
  
```{r}
TSRFunc <- function(x){
  data.frame(value = x[abs(x-mean(x))>3*sd(x)],index = which(abs(x-mean(x))>3*sd(x)))
}
data <- iris$Sepal.Width
TSRFunc(data)
```

**#2** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the 1.5*IQR Rule given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers.

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

```{r}
IQRFunc <- function(x){
 data.frame(value = x[x<fivenum(x)[2] - 1.5*IQR(x) | x>fivenum(x)[4] + 1.5*IQR(x)],index = which(x<fivenum(x)[2] - 1.5*IQR(x) | x>fivenum(x)[4] + 1.5*IQR(x)))
}
data <- iris$Sepal.Width
IQRFunc(data)
```

**#3** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization) Create a custom function that detects outliers based on the Hampel Identifier given a vector of data where `x <- iris$Sepal.Width`. Your function should return the values of the vector which are outliers.

- Hampel Identifier $|x-\tilde{x}|> 3 \cdot \tilde{\sigma}$  
      - where $\tilde{x}$ is the median of $x$  
      - where $\tilde{\sigma}=1.4826 \cdot \textrm{median}\{|x-\tilde{x}|\}$ is the median of the absolute deviation from the median (MADM or MAD) scale estimate

```{r}
HampelFunc <- function(x){
  data.frame(value = x[abs(x-median(x))>3*1.4826*median(abs(x-median(x)))],index = which(abs(x-median(x))>3*1.4826*median(abs(x-median(x)))))
}
data <- iris$Sepal.Width
HampelFunc(data)
```