---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 2 Practice Problems
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. (R Introduction, R Simple manipulations, Markdown) Suppose I don't remember exactly what R's median function does. Which line of code will query R about it and bring up the help page? Write your answer in bold and italic text.

a. `help(median)`
b. `?median`
c. both a and b
d. neither a nor b

***c. both a and b***

2. (Using Git and GitHub) Here is the scenario: 
```
The whole class is working on the final project Shiny app. Student X creates a menu for the final project Shiny app on their local machine. They show it to the Instructor and student X decides to commit and push their menu as a contribution to the final project. But Student X never pulled the class project repo first.
```
Will this result in a successful push or a merge conflict to the class project repo?

***merge conflict***

3. (R Simple manipulations, R Numbers and vectors) Calculate 10 × 4 - 2 × (4² ÷ 4) ÷ 2 ÷ 1/2 + 9 using R programming and a code chunk to show the result.

```{r}
10*4-2*((4^2)/4)/2/(1/2)+9
```