---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 7 Practice Problems
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#1** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution) Create a custom function that detects outliers based on the two outlier detection methods: Three Sigma Rule and 1.5*IQR Rule given a vector of data `x`. Your function should return the values of the vector which are outliers and the index location of the outlier (i.e. if the first value of x is an outlier then your function returns 1 and the value at x[1]).

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

Now, check to see if your outlier detector function works with the following as inputs:

i. `x <- rnorm(20)+runif(20)`

ii. `x <- ChickWeight$weight`

```{r}
outlier1 <- function(data){
    
  rbind(data.frame(value = data[abs(data-mean(data))>3*sd(data)],index = which(abs(data-mean(data))>3*sd(data))),
        data.frame(value = data[data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)],index = which(data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data))),
        data.frame(value = data[abs(data-median(data))>3*1.4826*median(abs(data-median(data)))],index = which(abs(data-median(data))>3*1.4826*median(abs(data-median(data))))))
}
x <- rnorm(20)+runif(20)
outlier1(x)
x <- ChickWeight$weight
outlier1(x)
```

**#2** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution) Create a custom function that detects outliers based on any of the two outlier detection methods: Three Sigma Rule and 1.5*IQR Rule given a vector of data `x` and one of the outlier detection methods as input. Your function should return the values of the vector which are outliers and the index location of the outlier (i.e. if the first value of x is an outlier then your function returns 1 and the value at x[1]).

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

Now, check to see if your outlier detector function works with the following as inputs:

i. `x <- rnorm(20)+runif(20)` and `Three Sigma Rule`

ii. `x <- ChickWeight$weight` and `1.5*IQR Rule`

```{r}
outlier2 <- function(data,method){
  
  if (method == "Three Sigma Rule"){
  data.frame(value = data[abs(data-mean(data))>3*sd(data)],index = which(abs(data-mean(data))>3*sd(data)))
  }
  else if (method == "1.5*IQR Rule"){data.frame(value = data[data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)],index = which(data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)))
  }
  else if (method == "Hampel Identifier"){
        data.frame(value = data[abs(data-median(data))>3*1.4826*median(abs(data-median(data)))],index = which(abs(data-median(data))>3*1.4826*median(abs(data-median(data)))))
  }
}
x <- rnorm(20)+runif(20)
outlier2(x, "Three Sigma Rule")
x <- ChickWeight$weight
outlier2(x, "1.5*IQR Rule")
```

**#3** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution, Apply family of functions) The code below is done with two `for` loops. Achieve the same result with apply family functions.

```{r}
set.seed(385)
nr <- nc <- 5
mat2 <- matrix(0, nrow = nr, ncol = nc)
for(i in 1:nr){ # rows loop
  for(j in 1:nc){ # columns loop
    z <- rnorm(1) # generate normal random deviate
    mat2[i, j] <- z # assign "large" value to matrix position (i,j)
  }
}
mat2
```

```{r}
set.seed(385)
nr <- nc <- 5
mat2 <- matrix(0, nrow = nr, ncol = nc)
func <- function(a){a <- rnorm(1)}
t(apply(mat2,c(1,2),func))
```
