---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 7 Practice Problems Solutions
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#1** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution) Create a custom function that detects outliers based on the three outlier detection methods: Three Sigma Rule, 1.5*IQR Rule, and Hampel Identifier given a vector of data `x`. Your function should return the values of the vector which are outliers and the index location of the outlier (i.e. if the first value of x is an outlier then your function returns 1 and the value at x[1]).

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

- Hampel Identifier $|x-\tilde{x}|> 3 \cdot \tilde{\sigma}$  
      - where $\tilde{x}$ is the median of $x$  
      - where $\tilde{\sigma}=1.4826 \cdot \textrm{median}\{|x-\tilde{x}|\}$ is the median of the absolute deviation from the median (MADM or MAD) scale estimate

Now, check to see if your outlier detector function works with the following as inputs:

i. `x <- rnorm(20)+runif(20)`

ii. `x <- ChickWeight$weight`

```{r}
outlier_detector1 <- function(data){
  rbind(data.frame(outlier_value=data[abs(data-mean(data))>3*sd(data)], outlier_position=which(abs(data-mean(data))>3*sd(data))), 
        data.frame(outlier_value=data[data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)], outlier_position=which(data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data))),
         data.frame(outlier_value=data[abs(data-median(data))>3*1.4826*median(abs(data-median(data)))], outlier_position=which(abs(data-median(data))>3*1.4826*median(abs(data-median(data))))))
}
```

```{r}
x <- rnorm(20)+runif(20)
outlier_detector1(x)

x <- ChickWeight$weight
outlier_detector1(x)
```


**#2** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution) Create a custom function that detects outliers based on any of the three outlier detection methods: Three Sigma Rule, 1.5*IQR Rule, and Hampel Identifier given a vector of data `x` and one of the outlier detection methods as input. Your function should return the values of the vector which are outliers and the index location of the outlier (i.e. if the first value of x is an outlier then your function returns 1 and the value at x[1]).

- Three Sigma Rule $|x-\bar{x}|>3 \cdot \hat{\sigma}$ 

- 1.5*IQR Rule $x<Q_1-1.5\cdot IQR$ or $x>Q_3+1.5 \cdot IQR$

- Hampel Identifier $|x-\tilde{x}|> 3 \cdot \tilde{\sigma}$  
      - where $\tilde{x}$ is the median of $x$  
      - where $\tilde{\sigma}=1.4826 \cdot \textrm{median}\{|x-\tilde{x}|\}$ is the median of the absolute deviation from the median (MADM or MAD) scale estimate

Now, check to see if your outlier detector function works with the following as inputs:

i. `x <- rnorm(20)+runif(20)` and `Three Sigma Rule`

ii. `x <- ChickWeight$weight` and `1.5*IQR Rule`

iii. `x <- iris$Sepal.Length` and `Hampel Identifier`

```{r}
outlier_detector2 <- function(data, method){
  if (method == "Three Sigma Rule")
    data.frame(outlier_value=data[abs(data-mean(data))>3*sd(data)], outlier_position=which(abs(data-mean(data))>3*sd(data)))
  else if (method == "1.5*IQR Rule")
    data.frame(outlier_value=data[data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data)], outlier_position=which(data<fivenum(data)[2]-1.5*IQR(data) | data>fivenum(data)[4]+1.5*IQR(data))) 
  else if (method == "Hampel Identifier")
    data.frame(outlier_value=data[abs(data-median(data))>3*1.4826*median(abs(data-median(data)))], outlier_position=which(abs(data-median(data))>3*1.4826*median(abs(data-median(data)))))
}
```

```{r}
x <- rnorm(20)+runif(20)
outlier_detector2(x, "Three Sigma Rule")

x <- ChickWeight$weight
outlier_detector2(x, "1.5*IQR Rule")

x <- iris$Sepal.Length
outlier_detector2(x, "Hampel Identifier")
```


**#3** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions, R Lists and data frames, R Writing your own functions, Vectorization, R Grouping, looping, and conditional execution, Apply family of functions) The code below is done with two `for` loops. Achieve the same result with apply family functions.

```
set.seed(385)
nr <- nc <- 5
mat2 <- matrix(0, nrow = nr, ncol = nc)
for(i in 1:nr){ # rows loop
  for(j in 1:nc){ # columns loop
    z <- rnorm(1) # generate normal random deviate
    mat2[i, j] <- z # assign "large" value to matrix position (i,j)
  }
}
mat2
```

```{r}
set.seed(385)
nr <- nc <- 5
mat2 <- matrix(0, nrow = nr, ncol = nc)
t(apply(mat2, 1, function(a){
  rnorm(a)
}))
```


