---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 11 Practice Problems
### Created by Prof. Christopher Kinson

Using the sentence about the geese below, address the following in a using regular expression and string manipulation to show code and results of the code as answers to the questions:

```{r}
library(tidyverse)
sentence <- "Friends of the Geese are hosting a memorial service Saturday for the 175 geese killed this week by the Urbana Park District in its 'charity harvest.'"
```

a. Find the number of times the lower case letter "s" appears  

```{r}
str_count(sentence,"s")
```

b.  Count the number of times a word begins with the letter "t" 

```{r}
str_count(sentence,"\\b[T|t]\\w*\\b")
```

c. Find the number of times any digit between 1-9 appears

```{r}
str_count(sentence,"[1-9]")
```

d. Count the number of words that contain the letter "g" (case does not matter).

```{r}
str_count(sentence,"\\b\\w*[G|g]\\w*\\b")
```

e. Which words have consecutive characters appearing twice?

```{r}
str_extract_all(sentence,"\\b\\w*([A-Z|a-z])\\1\\w*\\b")
```

f. Show only the words excluding punctuation such that each word appears as a row in a matrix. This matrix should have 26 rows and 1 column.

```{r}
t(str_extract_all(sentence,"\\b\\w+\\b",simplify=TRUE))
```