---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 11 Practice Problems Solutions
### Created by Prof. Christopher Kinson

Using the sentence about the geese below, address the following in a using regular expression and string manipulation to show code and results of the code as answers to the questions:

```{r}
library(tidyverse)
sentence <- "Friends of the Geese are hosting a memorial service Saturday for the 175 geese killed this week by the Urbana Park District in its 'charity harvest.'"
```

a. Find the number of times the lower case letter "s" appears  
```{r}
str_count(sentence, "s")
```

b.  Count the number of times a word begins with the letter "t" 
```{r}
str_count(sentence, "th+")
#sum(str_count(str_extract_all(sentence, boundary("word"))[[1]], "^t+")) # also works
```

c. Find the number of times any digit between 1-9 appears
```{r}
str_count(sentence, "[1-9]")
```
  
d. Count the number of words that contain the letter "g" (case does not matter).
```{r}
str_count(sentence, "g|G")
```
  
e. Which words have consecutive characters appearing twice?
```{r}
str_extract_all(sentence, "\\w+[el]{2}\\w+") #only works in this specific case, but not very general
str_extract_all(sentence, "\\w+([a-z])\\1\\w+") #works more generally
```

f. Show only the words excluding punctuation such that each word appears as a row in a matrix. This matrix should have 26 rows and 1 column.
```{r}
t(str_extract_all(sentence, boundary("word"), simplify=TRUE))
```

