---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Week 13 Notes
### Created by Christopher Kinson


***


## Things Covered in this Week's Notes

- Shiny app development Part 1


***


## We're going to build Shiny apps!!!

Execute the chunks within this Rmd file on your local computer to see and interact with the Shiny app.

```{r, shinyapp1, eval=FALSE}
#
# This is a Shiny web application. You can run the application by clicking
# the 'Run App' button above.
#
# Find out more about building applications with Shiny here:
#
#    http://shiny.rstudio.com/
#

library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
  
    # Sidebar with a slider input for number of bins 
      sliderInput(inputId = "bins",
                        label = "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30), #the comma is important
      
    # Show a plot of the generated distribution
      plotOutput("distPlot")
      )

# Define server logic required to draw a histogram
server <- function(input, output) {

    output$distPlot <- renderPlot({
        # generate bins based on input$bins from ui.R
        x    <- faithful[, 2]
        bins <- seq(min(x), max(x), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
  })
}

# Run the application 
shinyApp(ui = ui, server = server, options = list(height = 500))

```

## What are shiny apps? Why should we use them?

These are free, easy-to-use, almost-easy-to-create, interactive data products. Shiny apps represent digestible content for anyone to learn about statistics and data through direct manipulation of the data values. The nice thing is that the app is **reactive** and changes with every confirmed change the user (of the app) makes.


## How to build a Shiny app? 

There are some basic functions and syntax that is common to all shiny apps, while other functions and tasks will be specific to the type of data product you want to show.

## Please watch Parts 1-3 of the [Shiny app tutorial](https://shiny.rstudio.com/tutorial/) by RStudio to learn more on your own. 

The following notes are from Grolemund's video and slides https://github.com/rstudio-education/shiny.rstudio.com-tutorial.

Let's begin by looking at the coding for the above watered down version of the Faithful app (the template is in RStudio already).

```{r, shinyapp2, eval=FALSE}
#
# This is a Shiny web application. You can run the application by clicking
# the 'Run App' button above.
#
# Find out more about building applications with Shiny here:
#
#    http://shiny.rstudio.com/
#

library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
  
    # Sidebar with a slider input for number of bins 
      sliderInput(inputId = "bins",
                        label = "Number of bins:",
                        min = 1,
                        max = 50,
                        value = 30), #the comma is important
      
    # Show a plot of the generated distribution
      plotOutput("distPlot")
      )

# Define server logic required to draw a histogram
server <- function(input, output) {

    output$distPlot <- renderPlot({
        # generate bins based on input$bins from ui.R
        x    <- faithful[, 2]
        bins <- seq(min(x), max(x), length.out = input$bins + 1)

        # draw the histogram with the specified number of bins
        hist(x, breaks = bins, col = 'darkgray', border = 'white')
    })
}

# Run the application 
shinyApp(ui = ui, server = server)

```


Every Shiny app must have 3 components:  
1. **ui** or user interface  
2. **server** function (followed by grouped expression`{}`)  
3. **shinyApp** function

### ui

The user interface contains code that builds the web document by creating html for the app. This ui object contains the `fluidPage()` function which contains all the "layers" of the design of the app. The app will need **input** and **output** designs.

There are several **input** designs:  

- buttons `actionButton()` or `submitButton()`

- single or group checkboxes `checkboxInput()` or `checkboxGroupInput()` 

- date input `dateInput()` or range `dateRangeInput()`  

- file input `fileInput()`  

- numeric input `numericInput()` 

- text input `textInput()` 

- radio buttons `radioButtons()`   

- select box (i.e. dropdown menu) `selectInput()`

- sliders `sliderInput()`   

For these input designs we need an `inputId` for identifying the input (and for use with the server function) and a `label` or explanation of the input type. Users will read this label.

There are several **output** designs:

- interactive table `dataTableOutput()`

- raw html `htmlOutput()`

- image `imageOutput()`

- plot `plotOutput()`

- table `tableOutput()`

- text `textOutput()` or `verbatimTextOutput()`

- a Shiny UI element `uiOutput()`

For these output designs, we need an `outputId`. None of the fluidPage arguments will create any meaningful output for us to see. It does create html for the page.


### server(){}

The `server()` function is the R code (that we're used to) giving instructions for the inputs and outputs. The `server()` function creates/builds/re-builds the **output** of the app such as the plot, table, and text by first recognizing the **input** (the data values). The `server()` function must have both **input** and **output**.

The **input** in the `server()` function must be named as `input$`* where the asterisk represents the naming of the object which should match the `inputId` (created in the ui portion of the code). the input is a list.

If building an **output** object, it needs to be saved as `output$`* where the asterisk represents the naming of the object which should match the `outputId` (created in the ui portion of the code). The output is a list.

To display the output object, we use the `render*()` function, where the asterisk is a particular type of render function. Usually the `render*()` function has a corresponding output type (see ui above).

- interactive table `renderDataTable()`

- image `renderImage()`

- plot `renderPlot()`

- a code block `renderPrint()`

- table `renderTable()`

- character string `renderText()`

- a shiny UI element `renderUI()`

#### Reactivity

When a user changes an input value in the app and the app output changes as a result, that is called **reactivity**.

If we want the **output** (histogram in the above app) to change after the user changes the **input** (slider in the above app), then we need to use the `input$`* matching the Id of the input in the ui and that input must go inside the `render*()` function. *If we do this correctly, then the reactivity occurs automatically!*

**Reactive values** (i.e. the `input$`) work together with **reactive functions.** We'll talk about some reactive functions on Wednesday and Monday of next week. Friday of this week we will tweak more shiny app examples, address the final project tasks and homework 5.


### shinyApp()

The `shinyApp()` function serves as the "knitting" function to weave together the ui and `server()` function. This function creates the app which is running locally on your computer.


### Task to Do

Here are more examples of Shiny apps:  
- https://shiny.rstudio.com/gallery/kmeans-example.html  
- https://shiny.rstudio.com/gallery/faithful.html  
- https://shiny.rstudio.com/gallery/movie-explorer.html  

1. Using Shiny, make a histogram with the `faithful` data (which is already inside of R). Allow the user to change the number of observations for the waiting column.

```{r}
library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "obs", label = "Choose number of observations for waiting", value = 50, min = 1, max = 272),
  plotOutput("hist")
)

server <- function(input,output) {
  output$hist <- renderPlot({hist(head(faithful$waiting,input$obs))})
}

shinyApp(ui=ui,server=server)
```

2. Using Shiny, make a histogram with the `faithful` data (which is already inside of R). Allow the user to change the number of observations for the eruptions column.

```{r}
library(shiny)

ui <- fluidPage(
    sliderInput(inputId = "obs", label = "Choose number of observations for eruption", value = 50, min = 1, max = 272),
    plotOutput("hist")
)

server <- function(input,output) {
    output$hist <- renderPlot({hist(head(faithful$eruption,input$obs))})
}

shinyApp(ui=ui,server=server)
```

3. Using Shiny, make a scatter plot with the `faithful` data (which is already inside of R). Allow the user to change the number of observations for both x (eruptions) and y (waiting) variables.

```{r}
library(shiny)

ui <- fluidPage(
    sliderInput(inputId = "obs", label = "Choose number of observations for both eruption and waiting", value = 50, min = 1, max = 272),
    plotOutput("scatter")
)

server <- function(input,output) {
    output$scatter <- renderPlot({plot(head(faithful$eruption,input$obs),head(faithful$waiting,input$obs))})
}

shinyApp(ui=ui,server=server)
```