---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 6 Notes
### Created by Christopher Kinson


***


## Things Covered in this Week's Notes

- Vectorization

- **An Introduction to R** *Chapter 10 Writing your own functions*

- Helpful coding principles

***

## Vectorization

Vectorization is the idea of taking advantage of the vector structure in your programming software to do repeatable computations efficiently. It is a concept that will surely speed up computing time (compared to traditional loops) and works to optimize repeated operations. Vectorized functions occur in two different ways. 

1. a function that does the same operation on a vector of values as it would do on each single value and returns a vector of values e.g. `sqrt()`

2. a function that takes a vector and produces one single value based on the whole vector e.g. `sum()`

For more information on vectorization in R, do read **[The R Inferno](https://www.burns-stat.com/pages/Tutor/R_inferno.pdf)**, particularly Chapters 3-4.

Here are some vectorized functions in R: `ifelse()`, `is.na()`, `sqrt()`, `log()`, `abs()`, `sum()`, `colSums()`, `rowSums()`, `mean()`, `colMeans()`, `rowMeans()`, `unique()`. Check the table below. Some are more popular than others. You may find other functions that are even more useful.

function | sample code | part of output |
---|---|---|
length | `length(iris$Species)` | `r length(iris$Species)` |
nrow | `nrow(iris)` | `r nrow(iris)` |
ncol | `ncol(iris)` | `r ncol(iris)` |  
seq_along | `seq_along(c(1, 3, 5, 7, 8, 9))` | `r seq_along(c(1, 3, 5, 7, 8, 9))` |
ifelse | `ifelse(iris$Species=="setosa",1,0)`| `r head(ifelse(iris$Species=="setosa",1,0))`|
paste | `paste(iris$Species,"iris",1:150)`| `r head(paste(iris$Species,"iris",1:150))`| 
split | `split(iris,iris$Species)`| |
which | `which(iris$Sepal.Length>6.8)`| `r which(iris$Sepal.Length>6.8)`|
rowSums | `rowSums(iris[,-5])` | `r head(rowSums(iris[,-5]))`|
colSums | `colSums(iris[,-5])` | `r colSums(iris[,-5])`|
rowMeans | `owMeans(iris[,-5])` | `r head(rowMeans(iris[,-5]))`|
colMeans | `colMeans(iris[,-5])` | `r head(colMeans(iris[,-5]))`|
complete.cases| `complete.cases(iris$Species)` | `r head(complete.cases(iris$Species))`|
is.na | `is.na(iris)` | `r head(is.na(iris))`|

The code below shows an advantage in coding efficiency and readability through the use of vectorization.

```
#not vectorized
set.seed(13)
n <- 10
zz<-rep("",n)
for (i in 1:n){
 z <- rnorm(1) # generate normal random variable
 if (abs(z) > 1) zz[i]="Yes!"
}
zz

#vectorized
set.seed(13)
z<-rnorm(10)
zz[abs(z)>1]<-"Yes!"
zz
```

```{r}
1+2+3+4+5+6 #not vectorized

sum(1:6) #vectorized
```

```{r}
# determine if 7919 is prime or not
7919%%3 #not a vectorized approach
7919%%5
7919%%7
7919%%11
7919%%13 #...
7919%%89

# determine if 7919 is prime or not
num <- 7919%%(1:7919) #a vectorized approach
num[num==0]
#it's prime
```

***


## **An Introduction to R** *Chapter 10 Writing your own functions*

The crux of what you do as a programmer is create your own tools. These tools can greatly speed up data analysis, data wrangling, data visualization, modeling, etc. R allows users to create their own functions (i.e. user-defined functions) using the `function` and with at least one argument that can be named or unnamed:

```
newfunction <- function(argument1, namedarg=argument2){
  expressions
}
```

Using the function to see the results becomes:

```
newfunction(x,y)
```

Your custom functions can be complicated or simple. It will depend on the nature of what you want to build and your programming skill set.

### Example 1: A t-statistic for the two-sample case (found in the Intro to R textbook)

```{r}
twosam <- function(y1, y2) {
  n1 <- length(y1); n2 <- length(y2)
  yb1 <- mean(y1); yb2 <- mean(y2)
  s1 <- var(y1); s2 <- var(y2)
  s <- ((n1-1)*s1 + (n2-1)*s2)/(n1+n2-2)
  tst <- (yb1 - yb2)/sqrt(s*(1/n1 + 1/n2))
  tst
}
twosam(rnorm(20),rnorm(20))
```


### Example 2: An n-dimensional correlation matrix with exchangeable correlation rho

```{r}
exc <- function(n,rho){
  mat <- array(rho,dim=c(n,n))
  mat2 <- array(c(1:n,1:n),dim=c(n,2))
  mat[mat2] <- 1 
  mat
}
exc(5,0.5)
```


### Example 3: Aggregating over a dataset to compute the MAD (median absolute deviation from the median)

```{r}
madm <- function(x){
  median(abs(x-median(x)))
}
madm(rnorm(2,20))
```


***

## Helpful coding principles

- DRY. Don't Repeat Yourself.

- KISS. Keep It Simple Shawty.

- When it is is reasonable (time & effort) do vectorize your code. Do not vectorize your code when it is not reasonable (time & effort).

- Document your code with comments.
