---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 1 Practice Problems
### Created by Christopher Kinson

1. (Syllabus, Markdown) Who are the teaching assistant (TA) and course assistants (CAs)? Write your answer as a list with the TA's name in bold text and each CA's name in italic text.

*The TA is Michael Wieck-Sosa. As of 08-22-2020, there are no CAs for the course.*

2. (Syllabus, GitHub, Markdown) Let the following list represent the incorrect order of steps to begin and submit the first homework assignment. Now, list the items in the correct order.

The correct order is:

**1** Pull homework1 from the homework repo in GitHub.

**2** Start working on the homework1 by myself.

**3** When I get stuck on a problem in homework1, ask a question in the Issues page.

**4** Push homework1 into my repo (the one named with my netid).


3. (RMarkdown) Using RMarkdown, place the following code in an R code chunk such that it runs upon rendering (or knitting) this file to html.

```{r}
library(tidyverse)
fzpizza_data1 <- read_csv("https://uofi.box.com/shared/static/ko6uqp336nruywnormrp6euo9qrraigs.csv")
head(fzpizza_data1, 5)
tail(fzpizza_data1, 5)
``` 
