---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 4 Practice Problems Solutions
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**#1** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions) Using the vector X below
```{r, eval=FALSE}
set.seed(385)
X <- rnorm(30)
```

compute the following statistics for `X`:

**#a** median $med(x)$

**#b** absolute deviations from the median $y=|x-med(x)|$

**#c** median absolute deviations from the median $med(y)$ (i.e. the median of **part b**) 

```{r}
set.seed(385)
X <- rnorm(30)
median(X) #a
X-median(X) #b
median(X-median(X)) #c
```


**#2** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions) Let A and B be arrays
```{r, eval=FALSE}
A <- array((1:24)*(1/3),dim=c(2,3,4))
B <- array((1:24)*(3),dim=c(2,3,4))
```

**#a** Compute $C=3A + B$ and show C. The result is of dimension (2,3,4).

**#b** Compute $D=AB$ and show D. The result is a single value.

**#c** Use the `cbind` function with a vector and arithmetic to create an array that is the same as `B[,,1]`.
  
```{r}
A <- array((1:24)*(1/3),dim=c(2,3,4))
B <- array((1:24)*(3),dim=c(2,3,4))
C <- 3*A+B
C #a
D <- A%*%B
D #b
B[,,1]
cbind(c(3,6),c(3,6)+6,c(3,6)+12) #c
```


**#3** (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement, R Arrays and matrices, R Probability distributions) Let A and B be arrays
```{r, eval=FALSE}
A2 <- array((1:24)*(1/3),dim=c(2,3))
B2 <- array((1:24)*(3),dim=c(2,3))
```

**#a** Compute `C2=3*A2 + B2` and show C2. The result is of dimension (2,3).
  
**#b** Compute `D2=2*A2 %*% t(B2)` and show D2. The result is of dimension (2,2).
  
**#c** Compute `outer(D2,D2, "*")`. The result is of dimension (2,2,2,2).

```{r}
A2 <- array((1:24)*(1/3),dim=c(2,3))
B2 <- array((1:24)*(3),dim=c(2,3))
C2=3*A2 + B2
C2
D2=2*A2 %*% t(B2)
D2
outer(D2,D2+2, "+")
```

