---
title: "STAT 385 Statistics Programming Methods - Fall 2020"
output: html_document
---
## Week 3 Practice Problems
### Created by Christopher Kinson

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement) Create a vector (named as `vec1`) with 11 elements from a equally spaced sequence of that begins with 1 and ends with 21. Print the resulting `vec1`.

```{r}
vec1 <- seq(1,21,2)
vec1
```

2. (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement) Create a vector (named as `vec2`) which coerces vec1 to a factor with levels being the corresponding letters of the alphbet at the positions of the values of `vec1`. Print the resulting `vec2`.

```{r}
vec2 <- factor(vec1)
levels(vec2) <- letters[vec1]
vec2
```

3. (RMarkdown, R Objects, their modes, and attributes, R Simple manipulations and Numbers and vectors, Object arrangement) Create a vector (named as `vec3`) which sorts `vec2` in decreasing order. Print the resulting `vec3`.

```{r}
vec3 <- sort(vec2,decreasing = TRUE)
vec3
```